Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-08-06 Report/kkt_mse_approach.tex
kkt corrections

In order to perform the distributed approach, we can consider the method proposed in \cite{komulainen2013effective}, where the users evaluate the \ac{MSE}, and the dual variable \me{\alpha_{l,k,n}} using the transmission made by all \acp{BS} using the transmit precoders \me{\mvec{m}{l,k,n}^{(i-1)}}. Once the dual variables are evaluated, the users will notify the dual variables and the receive beamformers to the \acp{BS} using the uplink precoded pilot signaling, where the precoders used for the uplink pilots are the receive beamformers \me{\mvec{\tilde{w}}{l,k,n}^{(i-1)} = \sqrt{\alpha_{l,k,n}^{(i-1)}}\mvec{w}{l,k,n}^{\ast(i-1)}} evaluated at the receivers. Upon receiving the uplink precoded pilots by the \ac{BS}, the effective channel \me{\mvec{H}{b,k,n}^{\mathrm{T}} \mvec{\tilde{w}}{l,k,n}^{(i-1)}} can be used in the expression \eqref{kkt-mse-4.3} to update the transmit precoders at the respective \acp{BS}, \me{\mbf{x}^{\ast}} represents the conjugate of \me{\mbf{x}}. The algorithmic representation of the distributed \ac{MSE}-\ac{KKT} scheme is shown in Algorithm. \ref{algo-4}. In this approach, the coordinating \acp{BS} are required to perform short transmissions to update the transmit and the receive beamformers before the actual data transmission.
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{C}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=1} and the receive beamformers \me{\mbf{w}^{(0)}_{l,k,n}} randomly}
 \KwInit{\me{\epsilon_{l,k,n}^{(0)}} randomly and the dual variables \me{{\alpha}_{l,k,n}^{(0)} = 1}}
 set the maximum iteration counter \me{I_{\max},J_{\max}} to a valid number. \;
 \Repeat{until convergence or \me{j \geq J_{\max}}}{
 \lForEach{\ac{BS} \me{b \in \mc{B}}}
 \Repeat{until convergence or \me{i \geq I_{\max}}}{
	update the transmit precoders \me{\mvec{M}{k,n}^{(i)}} using \eqref{kkt-mse-4.3}, where \me{\delta_b} is identified by the bisection method satisfying \eqref{kkt-mse-1.4}. \;
    update \me{\epsilon_{l,k,n}^{(i)}}, \me{t_{l,k,n}^{(i)}}, \me{\sigma_{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} using \eqref{kkt-mse-4.4} and \eqref{kkt-mse-4.5}, \eqref{kkt-mse-4.2} and \eqref{kkt-mse-4.1} at the receivers using the transmit precoders \me{\mvec{M}{k,n}^{(i)}} from all \acp{BS}. \;
    perform precoded uplink pilot transmission from all users with the precoder \me{\mvec{\tilde{w}}{l,k,n}^{(j)}}. \;
    $i = i + 1.$ \;
  }
 evaluate the receive beamforming vector \me{\mvec{W}{k,n}^{(j+1)}} using \eqref{kkt-mse-4.6}. \;
 inform \acp{BS} using uplink precoded pilot transmission from all users with the precoder \me{\mvec{\tilde{w}}{l,k,n}^{(j+1)}}. \;
 \me{j = j + 1}. \;
 }
 \caption{\ac{KKT} approach for the \ac{JSFRA} scheme}
 \label{algo-4}
\end{algorithm}

>>>>

In order to perform the distributed approach, we can consider the method proposed in \cite{komulainen2013effective}, where the users evaluate the \ac{MSE}, and the dual variable \me{\alpha_{l,k,n}} using the transmission made by all \acp{BS} using the transmit precoders \me{\mvec{m}{l,k,n}^{(i-1)}}. Once the dual variables are evaluated, the users will notify the dual variables and the receive beamformers to the \acp{BS} using the uplink precoded pilot signaling, where the precoders used for the uplink pilots are the receive beamformers \me{\mvec{\tilde{w}}{l,k,n}^{(i-1)} = \sqrt{\alpha_{l,k,n}^{(i-1)}}\mvec{w}{l,k,n}^{\ast(i-1)}} evaluated at the receivers. Upon receiving the uplink precoded pilots by the \ac{BS}, the effective channel \me{\mvec{H}{b,k,n}^{\mathrm{T}} \mvec{\tilde{w}}{l,k,n}^{(i-1)}} can be used in the expression \eqref{kkt-mse-4.3} to update the transmit precoders at the respective \acp{BS}, \me{\mbf{x}^{\ast}} represents the conjugate of \me{\mbf{x}}. The algorithmic representation of the distributed \ac{MSE}-\ac{KKT} scheme is shown in Algorithm. \ref{algo-4}. In this approach, the coordinating \acp{BS} are required to perform short transmissions to update the transmit and the receive beamformers before the actual data transmission.
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{C}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=1} and the receive beamformers \me{\mbf{w}^{(0)}_{l,k,n}} randomly}
 \KwInit{\me{\epsilon_{l,k,n}^{(0)}} randomly and the dual variables \me{{\alpha}_{l,k,n}^{(0)} = 1}}
 for each \ac{BS} \me{b \in \mc{B}}, perform the following procedure \;
 set the maximum iteration counter \me{I_{\max},J_{\max}} to a valid number. \;
 initialize \me{j = 0}. \;
 \Repeat{until convergence or \me{j \geq J_{\max}}}{
 initialize \me{i = 0}. \;
 \Repeat{until convergence or \me{i \geq I_{\max}}}{
	update the transmit precoders \me{\mvec{M}{k,n}^{(i)}} using \eqref{kkt-mse-4.3}, where \me{\delta_b} is identified by the bisection method satisfying \eqref{kkt-mse-1.4}. \;
    update \me{\epsilon_{l,k,n}^{(i)}}, \me{t_{l,k,n}^{(i)}}, \me{\sigma_{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} using \eqref{kkt-mse-4.4} and \eqref{kkt-mse-4.5}, \eqref{kkt-mse-4.2} and \eqref{kkt-mse-4.1} at the receivers using the transmit precoders \me{\mvec{M}{k,n}^{(i)}} from all \acp{BS}. \;
    perform precoded uplink pilot transmission from all users with the precoder \me{\mvec{\tilde{w}}{l,k,n}^{(j)}}. \;
    \me{i = i + 1} \;
  }
 evaluate the receive beamforming vector \me{\mvec{W}{k,n}^{(j+1)}} using \eqref{kkt-mse-4.6}. \;
 inform \acp{BS} using uplink precoded pilot transmission from all users with the precoder \me{\mvec{\tilde{w}}{l,k,n}^{(j+1)}}. \;
 \me{j = j + 1}. \;
 }
 \caption{\ac{KKT} approach for the \ac{JSFRA} scheme}
 \label{algo-4}
\end{algorithm}

### 2014-08-06 Report/simulation_distributed.tex
kkt corrections

It can be seen from Fig. \ref{fig-d-3.2}, that the sum rate of the \me{q=1} \ac{KKT} approach outperforms other schemes in spite of Fig. \ref{fig-d-3.1} showing inferior queue minimizing performance. This due to the lack of the rate bounding constraint \eqref{kkt-mse-4.2}, which is simply given by \me{\sigma_{l,k,n}^{(i)} = \frac{a_k}{\log(2)}} providing no control for the excess allocations. For \me{q=2}, when the allocated rate exceeds the available queued bits, \me{\sigma_{l,k,n}^{(i)}} will be negative, thereby reducing \eqref{kkt-mse-4.1} as \me{\alpha_{l,k,n}^{(i)} < \alpha_{l,k,n}^{(i-1)}} to reduce the priority for the corresponding user in the next iteration. The performance of the closed form solution using the \ac{KKT} solution performs similar to the centralized schemes with \me{q=2} by limiting the rates beyond the available queued packets.
\begin{figure*}
\centering
\begin{subfigure}{0.49\textwidth}
\includegraphics[width=\textwidth]{fig-9}
\caption{Queue deviation}
\label{fig-d-3.1}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
\includegraphics[width=\textwidth]{fig-10}
\caption{Sum rate performance}
\label{fig-d-3.2}
\end{subfigure}
\caption{Convergence plot for \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 4,2,12,1 \rbrace} model}
\label{fig-d-3}
\end{figure*}

>>>>

It can be seen from Fig. \ref{fig-d-3.2}, that the sum rate of the \me{q=1} \ac{KKT} approach outperforms other schemes in spite of Fig. \ref{fig-d-3.1} showing inferior queue minimizing performance. This due to the lack of the rate bounding constraint \eqref{kkt-mse-4.2}, which is simply given by \me{\sigma_{l,k,n}^{(i)} = \frac{a_k}{\log(2)}}, providing no control for the excess allocations. For \me{q=2}, when the allocated rate exceeds the available queued bits, \me{\sigma_{l,k,n}^{(i)}} will be negative, thereby reducing \eqref{kkt-mse-4.1} as \me{\alpha_{l,k,n}^{(i)} < \alpha_{l,k,n}^{(i-1)}} to reduce the priority for the corresponding user in the next iteration. The performance of the closed form solution using the \ac{KKT} solution performs similar to the centralized schemes with \me{q=2} by limiting the rates beyond the available queued packets.
\begin{figure*}
\centering
\begin{subfigure}{0.49\textwidth}
\includegraphics[width=\textwidth]{fig-9}
\caption{Queue deviation}
\label{fig-d-3.1}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
\includegraphics[width=\textwidth]{fig-10}
\caption{Sum rate performance}
\label{fig-d-3.2}
\end{subfigure}
\caption{Convergence plot for \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 4,2,12,1 \rbrace} model}
\label{fig-d-3}
\end{figure*}
